---
title: "What did I do in 2023?"
subtitle: "Tracking my work and non-work time."
author: "Sam Zipper"
date: 2024-01-01
output: 
  html_document:
    toc: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(patchwork)

## ggplot theme
windowsFonts(Arial=windowsFont("TT Arial"))
theme_scz <- function(...){
  theme_bw(base_size=10, base_family="Arial") + 
    theme(
      text=element_text(color="black"),
      plot.title=element_text(face="bold", size=rel(1)),
      axis.title=element_text(face="bold", size=rel(1)),
      axis.text=element_text(size=rel(1)),
      strip.text=element_text(size=rel(1)),
      legend.title=element_text(face="bold", size=rel(1)),
      legend.text=element_text(size=rel(1)),
      panel.grid=element_blank(),
      plot.margin=unit(c(1,1,1,1), "mm"),
      strip.background=element_blank())
}

theme_set(theme_scz())

# color palette
col.cat.grn <- "#3cb44b"   # green
col.cat.yel <- "#ffe119"   # yellow
col.cat.org <- "#f58231"   # orange
col.cat.red <- "#e6194b"   # red
col.cat.blu <- "#0082c8"   # blue
col.gray <- "gray65"       # gray for annotation lines, etc

# colors by activity
pal_activity <-
  c("research" = "#e41a1c", 
    "logistics" = "#377eb8", 
    "talking" = "#4daf4a", 
"conference" = "#984ea3", 
"service" = "#ff7f00", 
"mentees" = "#ffff33",
"teaching" = "cyan",
"field" = "darkslategray")

# date limits for plots
date_limits <- as.POSIXct(c(ymd("2019-01-01"), ymd("2024-01-01")))

# load data
all_weeks_19 <- read_csv("https://raw.github.com/samzipper/SelfAnalysis/master/WorkTimeAnalysis_Hours_2019.csv", col_types = "ccccT")
all_weeks_20 <- read_csv("https://raw.github.com/samzipper/SelfAnalysis/master/WorkTimeAnalysis_Hours_2020.csv", col_types = "ccccT")
all_weeks_21 <- read_csv("https://raw.github.com/samzipper/SelfAnalysis/master/WorkTimeAnalysis_Hours_2021.csv", col_types = "ccccT")
all_weeks_22 <- read_csv("https://raw.github.com/samzipper/SelfAnalysis/master/WorkTimeAnalysis_Hours_2022.csv", col_types = "ccccT")
all_weeks_23 <- read_csv("https://raw.github.com/samzipper/SelfAnalysis/master/WorkTimeAnalysis_Hours_2023.csv", col_types = "ccccT")

# combine years
all_weeks <- 
  all_weeks_19 |> 
  dplyr::bind_rows(all_weeks_20) |> 
  dplyr::bind_rows(all_weeks_21) |> 
  dplyr::bind_rows(all_weeks_22) |> 
  dplyr::bind_rows(all_weeks_23) |> 
  subset(!is.na(Time))

# set factors for plotting
all_weeks$Day <- factor(all_weeks$Day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
all_weeks$Time <- factor(all_weeks$Time, levels = unique(all_weeks$Time))

# lump 'Travel' with 'Not Work' because if I was working on the plane I coded it as such
all_weeks$Activity[all_weeks$Activity == "travel"] <- "Not Work"

# summarize by week
weekly_hrs <- 
  all_weeks |> 
  dplyr::group_by(week_start_date) |> 
  dplyr::summarize(hours_worked = sum(Activity != "Not Work")*0.5) |> 
  dplyr::mutate(Year = lubridate::year(week_start_date))

# median and mode
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
median_hrs_wk <- median(weekly_hrs$hours_worked)
mode_hrs_wk <- getmode(weekly_hrs$hours_worked)

# summarize by year
weekly_hrs_yr <- 
  weekly_hrs |> 
  group_by(Year) |> 
  summarize(median_hrs_wk = median(hours_worked),
            mean_hrs_wk = mean(hours_worked),
            mode_hrs_wk = getmode(hours_worked))

# make a column for the actual date
day_to_number <- tibble(Day = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"),
                        DayNum = seq(0,6))
all_weeks <-  
  dplyr::left_join(all_weeks, day_to_number, by = "Day") |> 
  dplyr::mutate(date = week_start_date + days(DayNum))

# summarize by day
daily_hrs <- 
  all_weeks |> 
  dplyr::group_by(date, Day, week_start_date, Daytype) |> 
  dplyr::summarize(hours_worked = sum(Activity != "Not Work")*0.5) |> 
  dplyr::mutate(hours_worked_cut = cut(hours_worked, c(0, 0.25, 2, 6, 9, 24), include.lowest = T,
                                       labels = c("0", "0.5 - 2", "2.5 - 6", "6.5 - 9", "> 9.5")),
                Year = lubridate::year(date)) |> 
  subset(Year <= 2023)

daily_hrs$Day <- factor(daily_hrs$Day, levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

# figure out average hours worked per day of work
daily_hrs_yr <-
  daily_hrs |> 
  group_by(Year) |> 
  summarize(work_hours = sum(hours_worked),
            work_days = (sum(Daytype %in% c("conference", "work")) + sum(Daytype == "half day")*0.5),
            work_hrs.day = work_hours/work_days)

# figure out no work days per week
weekly_no_work <-
  daily_hrs |> 
  group_by(Year, week_start_date) |> 
  summarize(no_work_days = sum(hours_worked == 0))

yearly_no_work <-
  weekly_no_work |> 
  group_by(Year) |> 
  summarize(no_work_days_yr = sum(no_work_days),
            weeks_gt_0 = sum(no_work_days > 0),
            weeks_gt_1 = sum(no_work_days > 1))

# data frame with just work hours
all_work <- all_weeks[all_weeks$Activity != "Not Work", ]

# summarize by activity
work_by_activity <-
  all_work |> 
  dplyr::group_by(Activity) |> 
  dplyr::summarize(Hours = n()*0.5,
                   PrcTime = Hours/sum(weekly_hrs$hours_worked)) |> 
  dplyr::arrange(-Hours)

# yearly percent for each activity
work_by_year <-
  all_work |> 
  dplyr::mutate(Year = lubridate::year(week_start_date)) |> 
  dplyr::group_by(Year) |> 
  dplyr::summarize(halfhrs_yr = n(),
                   conference = sum(Activity=="conference")/halfhrs_yr,
                   logistics = sum(Activity=="logistics")/halfhrs_yr,
                   mentees = sum(Activity=="mentees")/halfhrs_yr,
                   field = sum(Activity=="field")/halfhrs_yr,
                   research = sum(Activity=="research")/halfhrs_yr,
                   service = sum(Activity=="service")/halfhrs_yr,
                   talking = sum(Activity=="talking")/halfhrs_yr,
                   teaching = sum(Activity=="teaching")/halfhrs_yr) |> 
  #dplyr::select(-halfhrs) |> 
  tidyr::pivot_longer(all_of(work_by_activity$Activity), names_to = "Activity", values_to = "Time_Prc") %>% 
  dplyr::mutate(Time_hrs = Time_Prc*halfhrs_yr*0.5)

# weekly percent for each activity
work_by_week <-
  all_work |> 
  dplyr::group_by(week_start_date) |> 
  dplyr::summarize(halfhrs = n(),
                   conference = sum(Activity=="conference")/halfhrs,
                   logistics = sum(Activity=="logistics")/halfhrs,
                   mentees = sum(Activity=="mentees")/halfhrs,
                   field = sum(Activity=="field")/halfhrs,
                   research = sum(Activity=="research")/halfhrs,
                   service = sum(Activity=="service")/halfhrs,
                   talking = sum(Activity=="talking")/halfhrs,
                   teaching = sum(Activity=="teaching")/halfhrs) |> 
  dplyr::select(-halfhrs) |> 
  tidyr::pivot_longer(all_of(work_by_activity$Activity), names_to = "Activity", values_to = "Time_Prc")

# set factor to order by most to least common
work_by_activity$Activity <- factor(work_by_activity$Activity, levels = work_by_activity$Activity)
all_work$Activity <- factor(all_work$Activity, levels = work_by_activity$Activity)
all_work$Activity <- factor(all_work$Activity, levels = work_by_activity$Activity)
work_by_week$Activity <- factor(work_by_week$Activity, levels = rev(work_by_activity$Activity))

# evening work, excluding field, conference, and travel
evening_work <-
  all_weeks |> 
  subset(Time %in% levels(all_weeks$Time)[23:35]) |> 
  subset(Daytype != "conference") %>% 
  subset(!(Activity %in% c("Not Work", "field", "conference", "travel"))) |> 
  mutate(Year = lubridate::year(week_start_date),
         Month = lubridate::month(week_start_date),
         YM = ymd(paste0(Year, "-", Month, "-01")))

evening_work_days <-
  evening_work |> 
  group_by(Year, week_start_date) |> 
  summarize(n_days = length(unique(Day))) |> 
  ungroup() |> 
  group_by(Year) |> 
  summarize(weeks_le_1 = 52-sum(n_days >  1))



```

# What is this?

This is the fourth edition of my time tracking analysis, adding in the 2023 calendar years. Past versions include:

 - [2019-2020 initial analysis](https://www.samzipper.com/post/2021-01-02_how-much-do-i-work/)
 - [2019-2021 follow-up](https://www.samzipper.com/post/2022-01-02-work-tracking-update/)
 - [2019-2022 follow-up](https://samzipper.com/post/2023-01-02-work-tracking-update.html)

Targets for 2023 included:

 - *Disconnecting:* 
    - Minimum of 1/day week with no work
    - Maximum of 1 day/week with evening work
 - *Time allocation:* 
    - Average 20% time research
    - <= 1 day/month field work.

The hours are tracked in a [simple google sheet](https://docs.google.com/spreadsheets/d/14ob3Ncsa_6ttzFWDB2FSn-NEEeSbKFZ2E7-P2qkWL48/edit#gid=244257672) at 30 min resolution and divided into a few simple categories:
 - `talking` = things that involved other people... meetings, seminars, zoom calls, etc.
 - `logistics` = stuff that is generally unimportant (but sometimes necessary). emails, expense reports, etc.
 - `research` = writing proposals or papers, making presentations, data analysis, reading papers, etc.
 - `mentees` = talking to or working with someone who I am mentoring such as a postdoc or student, including reviewing their work
 - `conference` = anything at a conference/workshop. I define it broadly so it includes going out for meals with colleagues, etc. - basically, any time I am in the presence of other people.
 - `service` = journal or grant reviews, committees, open science stuff, etc.

# Baseline stats: How much do I work?

My typical weekly workload remains consistent, with a median of `r weekly_hrs_yr$median_hrs_wk[weekly_hrs_yr$Year == 2023]` hours weekly:

```{r, echo=FALSE, fig.width=6, fig.height=4} 
ggplot(weekly_hrs, aes(x = week_start_date, y = hours_worked)) +
  geom_col(color = NA, aes(fill = factor(Year))) +
    scale_x_datetime(name = NULL, expand = c(0, 0), limits = date_limits) +
  scale_y_continuous(name = "Hours Worked in Week",
                     expand = expansion(mult = c(0, 0.05))) +
  geom_hline(yintercept = 40, color = col.gray) +
  scale_fill_viridis_d(guide = "none")
```

After account for time off by dividing total hours worked over number of work days, I've had a slight increase to `r round(daily_hrs_yr$work_hrs.day[daily_hrs_yr$Year == 2023], 1)` in 2023, up from `r round(daily_hrs_yr$work_hrs.day[daily_hrs_yr$Year == 2022], 1)` hours in 2022 but still down from my peak of `r round(daily_hrs_yr$work_hrs.day[daily_hrs_yr$Year == 2019], 1)` when I started tracking in 2019:

```{r, echo=FALSE, fig.width=4, fig.height=4} 
ggplot(daily_hrs_yr, aes(x = Year, y = work_hrs.day)) +
  geom_col(aes(fill = factor(Year))) +
  scale_y_continuous(name = "Total hours worked / work days",
                     breaks = seq(0, 9, 1),
                     expand = expansion(mult = c(0, 0.05))) +
  geom_hline(yintercept = 8, color = col.gray) +
  scale_fill_viridis_d(guide = "none")
```

```{r, echo=FALSE, fig.width=6, fig.height=8}
ggplot(daily_hrs, aes(x = Day, y = week_start_date, fill = hours_worked_cut)) +
  geom_raster() +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_datetime(name = "Week", expand = c(0, 0), limits = date_limits) +
  scale_fill_manual(name = "Daily Hours Worked",
                    values = c("gray65", "#ffffb2", "#fecc5c", "#fd8d3c", "#e31a1c")) +
  theme(legend.position = "bottom")
```

# Key goals/indicators

## 1. No work on days off

I set this goal at the beginning of 2021, after observing that I often was checking/responding to email on weekends. I've made continued improvements here each of the past three years, and am up to `r yearly_no_work$no_work_days_yr[yearly_no_work$Year == 2023]` days with absolutely no work in 2023.

```{r, echo=FALSE, fig.width=4, fig.height=4} 
ggplot(yearly_no_work, aes(x = Year, y = no_work_days_yr)) +
  geom_col(aes(fill = factor(Year))) +
  scale_y_continuous(name = "Days with 0 hrs work",
                     expand = expansion(mult = c(0, 0.05))) +
  scale_fill_viridis_d(guide = "none")
```


Looking weekly, this means there was at least one day completely off in `r yearly_no_work$weeks_gt_0[yearly_no_work$Year == 2023]` weeks. Since there are only 52 weeks in the year, minimal improvement is possible here, but I can aim for more weeks with at least two days completely off (i.e., both weekend days) - in 2023, I only had `r yearly_no_work$weeks_gt_1[yearly_no_work$Year == 2023]` weeks where I took two days off. This would require shifting my week planning from it's current timeslot (Sunday after lunch) to the end of the prior week (Friday afternoon). 

```{r, echo=FALSE, fig.width=8, fig.height=4} 
p_weeksoff1d <-
  ggplot(yearly_no_work, aes(x = Year, y = weeks_gt_0)) +
  geom_col(aes(fill = factor(Year))) +
  scale_y_continuous(name = "# of weeks with >= 1 day off",
                     expand = expansion(mult = c(0, 0.05))) +
  scale_fill_viridis_d(guide = "none")

p_weeksoff2d <-
  ggplot(yearly_no_work, aes(x = Year, y = weeks_gt_1)) +
  geom_col(aes(fill = factor(Year))) +
  scale_y_continuous(name = "# of weeks with >= 2 day off",
                     expand = expansion(mult = c(0, 0.05))) +
  scale_fill_viridis_d(guide = "none")

p_weeksoff1d + p_weeksoff2d +
  plot_layout(nrow = 1)
```

## 2. Do more research and less logistics

In 2023, I worked basically the exact same number of hours as 2022 (`r sum(work_by_year$Time_hrs[work_by_year$Year == 2023])` hours), with basically flat percent of my time allocated to logistics + talking (just over half, at `r round(100*sum(work_by_year$Time_Prc[work_by_year$Year == 2023 & work_by_year$Activity %in% c("logistics", "talking")]), 1)`%). However, my percent time for research did jump from `r round(100*sum(work_by_year$Time_Prc[work_by_year$Year == 2022 & work_by_year$Activity == "research"]), 1)`% in 2022 to `r round(100*sum(work_by_year$Time_Prc[work_by_year$Year == 2023 & work_by_year$Activity == "research"]), 1)`% in 2023, with a smaller jump for mentees (`r round(100*sum(work_by_year$Time_Prc[work_by_year$Year == 2022 & work_by_year$Activity == "mentees"]), 1)`% to `r round(100*sum(work_by_year$Time_Prc[work_by_year$Year == 2023 & work_by_year$Activity == "mentees"]), 1)`%). These increases were compensated for by minor decreases in field work, service, and teaching. So, trending in the right direction.

```{r, echo=FALSE, fig.width=6, fig.height=4}
all_work |> 
  dplyr::mutate(Year = lubridate::year(week_start_date)) |> 
  ggplot(aes(x = Year, fill = Activity)) +
  geom_bar() +
  scale_y_continuous(name = "Half-Hours",
                     expand = expansion(mult = c(0, 0.05))) +
  scale_fill_manual(values = pal_activity)
```


```{r, echo=FALSE, fig.width=4, fig.height=10}
work_by_year %>% 
  mutate(Activity_factor = factor(Activity, levels = levels(work_by_week$Activity))) %>% 
  ggplot(aes(x = Activity_factor, y = Time_hrs, fill = Activity)) +
  geom_col() +
  facet_wrap(~Year, ncol = 1, scales = "free_x") +
  scale_y_continuous(name = "Hours") +
  scale_fill_manual(values = pal_activity, guide = F) +
  theme(axis.text.x = element_text(angle = 25, hjust = 1, vjust = 1))
```

## 3. Less evening work

There remains a pretty clear seasonal pattern for evening work, with the most activity during the semester and less over the summer. In 2023, there were clear peaks in June (associated with a proposal deadline) and October/November (associated with teaching). 

```{r, echo=FALSE, fig.width=6, fig.height=8}
ggplot(subset(evening_work, Year >= 2020), aes(x = Month, fill = Activity)) +
  geom_bar() +
  facet_wrap(~Year, ncol = 1) +
  scale_y_continuous(name = "Half-Hours Worked After 6 PM",
                     expand = expansion(mult = c(0, 0.05))) +
  scale_x_continuous(name = "Month", breaks = seq(1,12)) +
  scale_fill_manual(values = pal_activity) +
  labs(title = "Monthly evening work") +
  geom_hline(yintercept = 10, color = col.gray)
```

In 2023, I was aiming for <= 1 day/week with evening work, as I thought that would be an easier goal to keep on top of than 10 hrs/week (my previous goal), and it looked like the shift in goal was effective. I'm up to `r evening_work_days$weeks_le_1[evening_work_days$Year == 2023]` weeks in 2023 with <= 1 day/week with evening work. So, while there hasn't been a big shift in terms of hours, it has been concentrated into a smaller number of days, leaving more evenings completely free.

```{r, echo=FALSE, fig.width=4, fig.height=4}
ggplot(evening_work_days, aes(x = Year, y = weeks_le_1)) +
  geom_col(aes(fill = factor(Year))) +
  scale_y_continuous(name = "Weeks with <= 1 day evening work",
                     expand = expansion(mult = c(0, 0.05))) +
  scale_fill_viridis_d(guide = "none")
```

## 4. Maximum 24 total field work days. 

```{r, include=FALSE}
n_field_days_yr <-
  all_work |> 
  subset(Activity == "field") |> 
  select(date) |> 
  unique() |> 
  mutate(Year = year(date)) |> 
  group_by(Year) |> 
  summarize(n_field_days = n())
```

The number of days I spent out in the field has continued to decrease from its high point in 2021, due primarily to a decrease in the field work requirements due to some projects ending and others reaching relatively stable conditions focused on field maintenance, rather than new field installations. 

```{r, echo=FALSE, fig.width=4, fig.height=4} 
ggplot(n_field_days_yr, aes(x = Year, y = n_field_days)) +
  geom_col(fill = "black") +
  scale_y_continuous(name = "Number of days with field work",
                     expand = expansion(mult = c(0, 0.05))) +
  geom_hline(yintercept = 24, color = col.gray)
```

# Any updates for 2024?

Most goals will remain the same:

  - Maintain 1/day week with no work and increase number of weeks with at least 2 days/week with no work.
  - Maximum of 1 day/week with evening work
  - Maintain research time at least 15% with target of 20%.
